Memories of migrations past: Sociality and cognition in dynamic, seasonal environments

### 2021-08-24 text/SocialMemoryMigration_FEE.tex
more ms finalization

		%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
	Seasonal migrations are a widespread and broadly successful strategy for animals to exploit periodic and localized resources over large spatial scales.  It remains an open and largely case-specific question whether long-distance migrations are resilient to environmental disruptions. High levels of mobility suggest an ability to shift ranges that can confer resilience. On the the other hand, a conservative, hard-wired commitment to a risky behavior can be costly if conditions change.

@@

		\tiny
		\keyFont{ \section{Keywords:} Sociality, seasonal migration, collective knowledge, diffusion-advection models, spatial memory} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
	\end{abstract}

@@

	For numerical runs with climate change, we first run a simulation with a given parameter set until stationarity, as above, and then begin shifting the location of the resource poleward with a slow, moderate or rapid trend ($\gamma_x = 0.25, 0.5$ and 1, respectively), and / or by adding stochasticity (spatial standard deviation 3, 6, 9 or 12). For stochasticity analyses, we compare foraging efficiency across a range of the reference memory parameter $\kappa$.  For analyses that included directional trends, with or without stochasticity, we quantified the ability of the system to keep track of climate change with a \emph{spatial adaptation} (SA) index. This index is the ratio of the slope of the memory-based migration location over time, i.e.~$\text{SA} = \widehat{\gamma}/\gamma_x$  where the adaptation slope estimate is the regression coefficient of the spatial coordinate of the migration against time (i.e.~$m_{x,i} = \widehat{\gamma_x} i + m_{x,0}$, where $i$ is the year), and $\gamma_x$ is the rate of drift of the resource peak (table \ref{table_definitions}). An SA equal to 1 suggests that the process is keeping up with climate change, an SA of 0 indicates that the process is not responding at all to climate change.  Values greater than 1 (super-adaptation) are possible, as are values less than 1, which correspond to a loss of migration behavior.

	All movement model parameters, resource parameters, and metrics are summarized in table \ref{table_definitions}.

@@

		\centering
		\begin{tabular}{ll|l}
			\hline\hline
			\multicolumn{3}{l}{\textbf{Memory migration model}}\\ 
			& $\varepsilon$ & Diffusion \\ 
			& $\alpha$ & Strength of resource following \\ 
			& $\beta$ & Strength of sociality \\ 
			& $\lambda$ & Spatial scale of sociality \\ 
			& $\kappa$ & Proportion of reference versus working memory \\ 
			& $x_1$, $x_2$ & location of population centroids in summer and winter \\ 
			& $t_1$, $\Delta t_1$ & start and duration of summer season \\ 
			& $t_2$, $\Delta t_2$ & start and duration of winter season \\ 
			& $\kappa$ & proportion memory allocated to reference \\
			&& (long-term) memory vs. working (short-term) memory \\ 
			\multicolumn{3}{l}{\textbf{Resource dynamics}} \\ 
			& $\tau$ & duration of period (year) \\ 
			& $m_x$, $-m_x$ & spatial coordinate of resource peak for summer and winter \\ 
			& $m_t$, $\tau - m_t$ & timing of resource peak for the summer and winter \\ 
			& $\sigma_x$, $\sigma_t$ & time duration and and spatial scale of resource pulse \\ 
			& $\gamma_x$, $\gamma_t$ & rate of change of peak location and timing of resource \\ 
			& $\psi_x$, $\psi_t$ & standard deviation of peak location and timing \\ 
			\multicolumn{3}{l}{\textbf{Metrics}}\\ 
			& $\text{MM}_x$ & spatial migration mismatch \\ 
			& $\text{MM}_t$ & temporal migration mismatch \\ 
			& $\text{TM}$ & total mismatch \\ 
			& $\text{FE}$ & foraging efficiency \\ 
			& $\text{SA}$ & spatial adaptation index \\ 
			\hline\hline
		\end{tabular}
	\end{table}

@@

	\begin{figure}
		\includegraphics[width = 0.9\textwidth]{figures/example1_adaptation.png}

@@

	\begin{figure}
		\includegraphics[width = \textwidth]{figures/StabilityResults.png} 
		\caption{\label{fig_phenologymatching} Migration phenology matching across six model parameters. Low and high diffusion ($\varepsilon = 1$ and $8$ in upper and lower panel blocks), tight, medium and loose swarms ($\lambda = 20, 50, 100$) left to right panels. Within each of these blocks, high values of the resource following parmaeters $\alpha$ from 0 to 1000 are left to right, and higher values of the sociality parameter $\beta$ are bottom to top. Within each of the combinations of $\varepsilon, \lambda, \alpha, \beta$, we show results ranging across 6 values of resource duration ($\sigma_t$, 3 to 15 left to right), and 6 values of resource extent ($\sigma_x$, 3 to 15 bottom to top), as in the zoomed-in panel (bottom right; drawn from $\alpha = 200$, $\beta = 200$, $\lambda = 100$). The color scheme reflects the total mismatch, i.e.~the sum of the absolute differences between the migration timing and locations from the resource peak.}
	\end{figure}

@@

	\begin{figure}
		\includegraphics[width = 0.9\textwidth]{figures/example2_learningtomigrate.png}

@@

	Overall, as expected, the greater the stochasticity, the lower the foraging efficiency.  Further, as we predicted, highest level of $\kappa$ can significantly help foraging efficiency, with some variation across the spatial scale of sociality, especially in more highly stochastic scenarios. When that scale of sociality is high enough ($\lambda = 120$, blue colors) there is greater probability of overlap with a stochastic resource, and a conservative, stable migratory regime is much more beneficial in the long run.

@@

		\caption{\label{fig_stochasticity} Foraging efficiency (FE) across various values of reference memory $\kappa$ (x-axis) for increasing amounts of interannual stochasticity ($\psi_x$, panels left to right), and two values of sociality spatial scale $\lambda = 80$ and $120$. For the processes with non-zero stochasticity ($\psi_x > 0$), the process was run 90 times for values of $\kappa$. Points represent the average of the FE's across all 50 years and 90 replicates, and shaded areas represent a locall regression (loess) smooth of FE against $\kappa$. In these scenarios, the resource following parameter $\alpha = 100$, the social attraction $\beta = 400$ and diffusion $\varepsilon = 4$.}

@@

	Animals navigate complex, dynamic and patchy environments. When there is a strongly localized and seasonal component to the resource dynamics, movement strategies limited to straightforward resource-following taxis necessarily fail to maximally exploit available resources. It is in these cases, quite common in the natural world, that seasonal migration becomes a viable, even necessary, strategy.  However, when resources start shifting in space and time -- as is occurring at an accelerated pace with recent global climate change -- the migration phenology itself must exhibit some plasticity. It is our conjecture that this plasticity is facilitated by a memory-driven process in which recent experiences inform strategic behaviors in subsequent years. The relatively simple, social and memory-driven mechanism we propose here was, in fact, able to adapt to long-term changes in resource dynamics, even with inter-annual stochasticity. The model thereby provides a framework with which the interaction  of memory, movement, social and resource dynamics can be further explored.

@@

	By allowing a population to adjust its migratory behavior based on recent experiences with the resource location, our model captured many fundamental phenomena related to migration. The model was able to emulate the successful navigation of an environment with temporally and spatially isolated seasonal resource patches, the emergence of a migratory behavior from an essentially resident or naive initial condition, and intrinsic robustness to changes in those environmental resources, whether those changes were steadily shifting trends or inter-annual stochasticity.  In a generalized way, it captured all three fundamental types of migration plasticity: the \textit{where}, the \textit{when} and the \textit{whether} \citep{Xu2021}.

	Our goal was to understand the combinations of factors that lead to a resilient migration behavior. We used foraging efficiency as a convenient metric of the utility of migration, though this was not a measure explicitly maximized by the model. Other metrics, such as foraging efficiency in a given season, or probability of survival or reproduction relative to resource availability \citep{Bauer2020} may respond differently across model parameters and could be useful in understanding the relative success of alternative migratory strategies in different contexts. However, the overall annually averaged metric provided the broadest linkage between resource dynamics and animals' locations and was consistent with the minimal biological assumptions and generality of our framework.

	Sociality parameters -- in particular, the spatial scale $\lambda$ -- were, unexpectedly among the most important parameters for determining the resiliency of the process. Populations with small spatial scales tended to have a more difficult time locking in to an adaptive migratory pattern, and only when social attraction was relatively weak. On the other hand, overly large spatial scales compromised the ability of the process to track climate change, due to a dilution of the population's ability to concentrate over available resource patches and remember the corresponding benefits.

@@

	Even with no strong intrinsic propensity to migrate and a weak phenological resource pulse to follow, our model captured the abililty to acquire a strong and adaptive migration behavior (figure \ref{fig_learningtomigrate}).  Learning migration, howewer, requires a very strong resource attraction, higher levels of exploratory behavior (e.g.~diffusion, and larger spatial scale of sociliaty), and -- often --- many more years, findings that echo empirical observations \citep{Jesmer2018}.

	Despite the ability of the process to adapt under many stable conditions, our migration model (and, perhaps, migration behaviors in general) can also be considered somewhat fragile. Under many shifting conditions, e.g.~increasing stochasticity, rapidly shifting resources, a shift in some of the system parameters, or even a shift in the spatial and temporal extent of resources, migration can collapse and turn into a non-migratory, residential behavior (figure \ref{fig_phenologymatching}). This sensitivity may explain why partially migratory populations are so common and, apparently, evolutionary stable \citep{Berthold1999, Chapman2011}, as well as the wide range of migration plasticity shown in wild populations, even within a species \citep{Xu2021}.

@@

%	Diffusion-advection models of animal movement and redistribution are grounded in the general idea that animal movements, somewhat like movements of physical particles, combine random (diffusive) components with directed (advective) components \citep{Skellam1951, Turchin1998, Okubo2001}. While direct relationships between diffusion models and movement data are somewhat tenuous \citep{Gurarie2011,Potts2020}, as a theoretical tool for exploring processes they are invaluable for their versatility and the relative ease of numeric computation of the partial differential equations (PDEs) that describe them mathematically. Refinements to diffusion-advection equations have revealed conditions under which non-local information gathering \citep{Fagan2017} and behavioral switching may confer foraging advantages \citep{Fagan2019}, in particular when resources are dynamic and patchy. While diffusion models that incorporate social movements (swarming, schooling, herds) have been well-studied \citep{Okubo1986, Gruenbaum1994, Mogilner1999}, the interacting role of memory and sociality in a specifically migratory framework has been relatively unexplored.

	All of the parameters in our model have well-defined biological interpretations. The diffusion ($\varepsilon$) captures short time-scaled randomness of movement, reflecting exploratory and short-term dispersive behavior. The foraging advection strength ($\alpha$) captures the attraction of the population to better quality resources at a relatively large scale. These two parameters, the basic ingredients in diffusion-advection models of animal movement, have direct parallels to empirically estimated properties of animal behavior: diffusion is closely related to families of random walk models \citep{Gurarie2011} while the advective taxis is related to the step- and resource selection functions that are routinely estimated from movement data \citep{Potts2020}. The spatial scale of the social group ($\lambda$) captures the spatial extent of the population, i.e.~a population-level home range \citep{Noonan2019}. 	We note, however, that diffusion-advection models can also be interpreted as a probabilistic description of a single individual's movement. In this case, $\lambda$ would correspond to an individual home-range and $\beta$ would be an individual's tendency to be drawn to the center of that home range, akin to an individual migratory Ornstein-Uhlenbeck process \citep{Gurarie2017}.

	The sociality parameter ($\beta$) quantifies the strength an individual's desire to approach the center of the social group. While this parameter is not typically measured, it may in principle be possible to estimate in a manner analogous to step-selection function by replacing environmental variables with presence of conspecifics as a covariate. The ratio between $\alpha$ and $\beta$ can be interpreted as the relative importance of foraging to social cohesion, which appears to be important in predicting the resilience of migration.

	Migration timing and location parameter can be straightforwardly estimated from movement data \citep{Cagnacci2015, Gurarie2019} and synchrony of migration timing and site fidelity are well-documented for many migratory species \citep{Joly2021}. Thus, for example, \citet{Gurarie2019} explicitly estimated the ranging area, timing, and seasonal range locations for migratory caribou, identifying the kind of inter-annual variation that is reflected in the stochastic scenarios explored here as well as trends in timing.

@@

	We'll note that our exploration of the model was not exhaustive. We did not explore, for example, the resilience of the migration process to changes in phenology timing. We hope that making the model available, including via an interactive interface, will facilitate those explorations.

@@

	Models have shown that collective knowledge is important, if not essential, to the evolution and process of migration \citep{Shaw2013, Guttal2010, Berdahl2018}. Many migratory organisms are social, and social learning is an acknowledged, non-genetic method for transmitting information \citep{Kashetsky}.  Because our model is not individual-based, we can not identify any specific mechanism (e.g.~leader-follower) of social information transfer.  But, in a generic way, our model assumes that migration is driven by a collective trigger for the timing and locations of seasonal ranges, which is consistent with the known social and exogenous (e.g.~daylength related) triggers for migration. Further, the underlying assumption of the migration ``urge'' is consistent with the strong endogenous programs to migrate, e.g.~the seasonal restlessness known as \emph{Zugunruhe}  exhibited by many birds \citep{Berthold1999, Helm2006}.

@@

	Given the slow scale of fitness selection and the constant change in environmental conditions, it is possible that certain inherent properties of populations, for example the ``conservatism'' captured by the $\kappa$ parameter, are themselves selected for to maximize resilience over a long time scale in stochastic environments.

@@

	\subsection{Drifting resource}
	The drifting resource function has the following properties

>>>>

	Seasonal migrations are a widespread and broadly successful strategy for animals to exploit periodic and localized resources over large spatial scales.  It remains an open and largely case-specific question whether long-distance migrations are resilient to environmental disruptions. High levels of mobility suggest an ability to shift ranges that can confer resilience. On the the other hand, a conservative, hard-wired commitment to a risky behavior can be costly if conditions change.

@@

		\tiny
		\keyFont{ \section{Keywords:} Sociality, seasonal migration, collective knowledge, diffusion-advection models, spatial memory} 
	\end{abstract}

@@

	For numerical runs with climate change, we first run a simulation with a given parameter set until stationarity, as above, and then begin shifting the location of the resource poleward with a slow, moderate or rapid trend ($\gamma_x = 0.25, 0.5$ and 1, respectively), and / or by adding stochasticity (spatial standard deviation 3, 6, 9 or 12). For stochasticity analyses, we compare foraging efficiency across a range of the reference memory parameter $\kappa$.  For analyses that included directional trends, with or without stochasticity, we quantified the ability of the system to keep track of climate change with a \emph{spatial adaptation} (SA) index. This index is the ratio of the slope of the memory-based migration location over time, i.e.~$\text{SA} = \widehat{\gamma}/\gamma_x$  where the adaptation slope estimate is the regression coefficient of the spatial coordinate of the migration against time (i.e.~$m_{x,i} = \widehat{\gamma_x} i + m_{x,0}$, where $i$ is the year), and $\gamma_x$ is the rate of drift of the resource peak (table \ref{table_definitions}). An SA equal to 1 suggests that the process is keeping up with climate change, an SA of 0 indicates that the process is not responding at all to climate change.  Values greater than 1 (super-adaptation) are possible, as are values less than 1, which correspond to a loss of migration behavior. All movement model parameters, resource parameters, and metrics are summarized in table \ref{table_definitions}. \\

@@

		\centering
		\begin{tabular}{ll|l}
			\hline\hline
			\multicolumn{3}{l}{\textbf{Memory migration model}}\\ 
			& $\varepsilon$ & Diffusion \\ 
			& $\alpha$ & Strength of resource following \\ 
			& $\beta$ & Strength of sociality \\ 
			& $\lambda$ & Spatial scale of sociality \\ 
			& $\kappa$ & Initial weighting of reference versus working memory \\ 
			& $x_1$, $x_2$ & location of population centroids in summer and winter \\ 
			& $t_1$, $\Delta t_1$ & start and duration of summer season \\ 
			& $t_2$, $\Delta t_2$ & start and duration of winter season \\ 
			&& (long-term) memory vs. working (short-term) memory \\ 
			\multicolumn{3}{l}{\textbf{Resource dynamics}} \\ 
			& $\tau$ & duration of period (year) \\ 
			& $m_x$, $-m_x$ & spatial coordinate of resource peak for summer and winter \\ 
			& $m_t$, $\tau - m_t$ & timing of resource peak for the summer and winter \\ 
			& $\sigma_x$, $\sigma_t$ & time duration and spatial scale of resource pulse \\ 
			& $\gamma_x$, $\gamma_t$ & rate of change of peak location and timing of resource \\ 
			& $\psi_x$, $\psi_t$ & standard deviation of peak location and timing \\ 
			\multicolumn{3}{l}{\textbf{Metrics}}\\ 
			& $\text{MM}_x$ & spatial migration mismatch \\ 
			& $\text{MM}_t$ & temporal migration mismatch \\ 
			& $\text{TM}$ & total mismatch \\ 
			& $\text{FE}$ & foraging efficiency \\ 
			& $\text{SA}$ & spatial adaptation index \\ 
			\hline\hline
		\end{tabular}
	\end{table}

@@

	\begin{figure}
		\includegraphics[width = 0.9\textwidth]{figures/example1_adaptation_lettered.png}

@@

	\begin{figure}
		\includegraphics[width = \textwidth]{figures/StabilityResults.png} 
		\caption{\label{fig_phenologymatching} Migration phenology matching across six model parameters. Low and high diffusion ($\varepsilon = 1$ and $8$ in upper and lower panel blocks), tight, medium and loose swarms ($\lambda = 20, 50, 100$) left to right panels. Within each of these blocks, high values of the resource following parmaeters $\alpha$ from 0 to 1000 are left to right, and higher values of the sociality parameter $\beta$ are bottom to top. Within each of the combinations of $\varepsilon, \lambda, \alpha, \beta$, we show results ranging across 5 values of resource duration ($\sigma_t$, 3 to 15 left to right), and 6 values of resource extent ($\sigma_x$, 3 to 15 bottom to top), as in the zoomed-in panel (bottom right; drawn from $\alpha = 200$, $\beta = 200$, $\lambda = 100$). The color scheme reflects the total mismatch, i.e.~the sum of the absolute differences between the migration timing and locations from the resource peak. The white squares represents parameter combinations where the PDE could not be solved for artifactual numerical reasons, that all correspond to a failed adaptation (high mismatch). }
	\end{figure}

@@

	\begin{figure}
		\includegraphics[width = 0.9\textwidth]{figures/example2_learningtomigrate_lettered.png}

@@

	Overall, as expected, the greater the stochasticity, the lower the foraging efficiency.  Further, as we predicted, highest level of $\kappa$ can significantly help foraging efficiency, with some variation across the spatial scale of sociality, especially in more highly stochastic scenarios. When that scale of sociality is high enough ($\lambda = 120$, blue colors) there is greater probability of overlap with a stochastic resource, and a conservative, stable migratory regime is much more beneficial in the long run. \\

@@

		\caption{\label{fig_stochasticity} Foraging efficiency (FE) across various values of reference memory $\kappa$ (x-axis) for increasing amounts of interannual stochasticity ($\psi_x$, panels left to right), and two values of sociality spatial scale $\lambda = 80$ and $120$. For the processes with non-zero stochasticity ($\psi_x > 0$), the process was run 90 times for values of $\kappa$. Points represent the average of the FE's across all 50 years and 90 replicates. In these scenarios, the resource following parameter $\alpha = 100$, the social attraction $\beta = 400$ and diffusion $\varepsilon = 4$. Note that the y-axis scale is constrained over a relatively narrow range (0.6-0.7). }

@@

	Animals navigate complex, dynamic and patchy environments. When there is a strongly localized and seasonal component to the resource dynamics, movement strategies limited to straightforward resource-following taxis necessarily fail to efficiently exploit available resources. It is in these cases, quite common in the natural world, that seasonal migration becomes a viable, even necessary, strategy.  However, when resources start shifting in space and time -- as is occurring at an accelerated pace with recent global climate change -- the migration phenology itself must exhibit some plasticity. It is our conjecture that this plasticity is facilitated by a memory-driven process in which recent experiences inform strategic behaviors in subsequent years.

	By allowing a population to adjust its migratory behavior based on recent experiences with the resource location, the model we presented here emulated (a) the successful navigation of an environment with temporally and spatially isolated seasonal resource patches, (b) the emergence of a migratory behavior from an essentially resident or naive initial condition, and (c) some intrinsic robustness to changes in those environmental resources, whether steadily shifting trends or inter-annual stochasticity. The relatively simple, social and memory-driven mechanism was able to adapt to long-term changes in resource dynamics, even with inter-annual stochasticity, and may thereby provide a framework with which the interaction of memory, movement, social and resource dynamics can be further explored.  % In a generalized way, it captured all three fundamental types of migration plasticity: the \textit{where}, the \textit{when} and the \textit{whether} \citep{Xu2021}.

	Importantly, our model was in no ways evolutionary, as it contained no birth-death processes or selection pressures. Thus, we used foraging efficiency as a convenient metric of the utility of migration, though this was not a measure explicitly maximized by the model. Other metrics, such as foraging efficiency in a given season, or probability of survival or reproduction relative to resource availability \citep{Bauer2020} may respond differently across model parameters and could be useful in understanding the relative success of alternative migratory strategies in different contexts. However, the overall annually averaged foraging efficiency metric provided the broadest linkage between resource dynamics and animals' locations and was consistent with the minimal biological assumptions and generality of our framework. \\

@@

	Our goal was to understand the combinations of factors that lead to a resilient migration behavior. The model we describe was a final iteration of a sequence of models which failed to develop or maintain social migration behavior. For example, in earlier versions memory was modeled as an attractive advection mathematically identical to the resource attraction, but with the attractor being the location of the population in previous years.  These models proved to be inefficient at generating a consistent social migratory behavior, i.e. only under very specific parameter combinations and ``easy'' conditions was a migratory equilibrium, and that equilibrium was highly unstable to perturbations.  Only a clear, directed advective process with an explicit seasonal signal (i.e.~the remembered migration timing, rates, and targets which were remembered in our model) could generate the patterns we aimed to capture.  This suggests, somewhat indirectly, that migration behavior is unique as a fundamental, long-term, and risky strategy, profoundly different from the kind of tactical resource response which governs shorter-scaled animal redistributions.

	Similarly, iterations of the model that did not have some amount of social cohesion tended to diffuse away without establishing a consistent, migratory stationary state. In fact, sociality parameters -- in particular, the spatial scale $\lambda$ -- were, unexpectedly among the most important parameters for determining the resiliency of the process. Populations with small spatial scales tended to have a more difficult time locking in to an adaptive migratory pattern, and only when social attraction was relatively weak. On the other hand, overly large spatial scales compromised the ability of the process to track climate change, due to a dilution of the population's ability to concentrate over available resource patches and remember the corresponding benefits.

@@

	Even with no strong intrinsic propensity to migrate and a weak phenological resource pulse to follow, our model captured the ability to acquire a strong and adaptive migration behavior (figure \ref{fig_learningtomigrate}).  Learning migration, howewer, requires a very strong resource attraction, higher levels of exploratory behavior (e.g.~diffusion, and larger spatial scale of sociliaty), and -- often --- many more years, findings that echo empirical observations \citep{Jesmer2018}.

	Despite the ability of the process to adapt under many stable conditions, our migration model (and, perhaps, migration behaviors in general) can also be considered somewhat fragile. Under many shifting conditions, e.g.~increasing stochasticity, rapidly shifting resources, a shift in some of the system parameters, or even a shift in the spatial and temporal extent of resources, migration can collapse and turn into a non-migratory, residential behavior (figure \ref{fig_phenologymatching}). This sensitivity may explain why partially migratory populations are so common and, apparently, evolutionary stable \citep{Berthold1999, Chapman2011}, as well as the wide range of migration plasticity shown in wild populations, even within a species \citep{Xu2021}.\\

@@

	Diffusion-advection models of animal movement and redistribution are grounded in the general idea that animal movements, somewhat like movements of physical particles, combine random (diffusive) components with directed (advective) components \citep{Skellam1951, Turchin1998, Okubo2001}. While direct relationships between diffusion models and movement data are somewhat tenuous \citep{Gurarie2011,Potts2020}, as a theoretical tool for exploring processes they are invaluable for their versatility and the relative ease of numeric computation of the partial differential equations (PDEs) that describe them mathematically. % Refinements to diffusion-advection equations have revealed conditions under which non-local information gathering \citep{Fagan2017} and behavioral switching may confer foraging advantages \citep{Fagan2019}, in particular when resources are dynamic and patchy. While diffusion models that incorporate social movements (swarming, schooling, herds) have been well-studied \citep{Okubo1986, Gruenbaum1994, Mogilner1999}, the interacting role of memory and sociality in a specifically migratory framework has been relatively unexplored.

	Despite its evident abstraction, our goal was to develop a model where all parameters have well-defined biological interpretations. The diffusion ($\varepsilon$) captures short time-scaled randomness of movement, reflecting exploratory and short-term dispersive behavior. The foraging advection strength ($\alpha$) captures the attraction of the population to better quality resources at a relatively large scale. These two parameters, the basic ingredients in diffusion-advection models of animal movement, have direct parallels to empirically estimated properties of animal behavior: diffusion is closely related to families of random walk models \citep{Gurarie2011} while the advective taxis is related to the step and resource selection functions that are routinely estimated from movement data \citep{Potts2020}. The spatial scale of the social group ($\lambda$) captures the spatial extent of the population, i.e.~a population-level home range \citep{Noonan2019}. We note, however, that diffusion-advection models can also be interpreted as a probabilistic description of a single individual's movement. In this case, $\lambda$ would correspond to an individual home-range and $\beta$ would be an individual's tendency to be drawn to the center of that home range, akin to an individual migratory Ornstein-Uhlenbeck process \citep{Gurarie2017}.

	The sociality parameter ($\beta$) quantifies the strength of an individual's desire to approach the center of the social group. While this parameter is not typically measured, it may in principle be possible to estimate in a manner analogous to step-selection function by replacing environmental variables with presence of conspecifics as a covariate. The ratio between $\alpha$ and $\beta$ can be interpreted as the relative importance of foraging to social cohesion, which appears to be important in predicting the resilience of migration.

	Migration timing, rate, and seasonal range location parameters can be straightforwardly estimated from movement data \citep{Cagnacci2015, Gurarie2019} and synchrony of migration timing and site fidelity are well-documented for many migratory species \citep{Joly2021}. Thus, for example, \citet{Gurarie2019} explicitly estimated the ranging area, timing, and seasonal range locations for migratory caribou, identifying the kind of inter-annual variation that is reflected in the stochastic scenarios explored here, as well as trends in timing.

@@

	Clearly, our exploration of the model was not exhaustive.  We did not explore, for example, the resilience of the migration process to changes in resource timing, which would correspond to the widely observed earlier onset of spring as measured by green-up and flowering phenology \citep{Cleland2007}. We hope that making the model available, including via the interactive interface, will facilitate further independent exploration of these processes.  \\

@@

	Models have shown that collective knowledge is important, if not essential, to the evolution and process of migration \citep{Shaw2013, Guttal2010, Berdahl2018}. Many migratory organisms are social, and social learning is an acknowledged, non-genetic method for transmitting information \citep{Kashetsky}. Furthermore, the general role of social learning for improving a population's ability to track resources has been studied not just in animal systems, but in synthetic systems inspired by social behavior of animals such as optimization heuristics algorithms and the study of swarm robotics \citep{Sahin2005, Brambilla2013}.  Because our model is not individual-based, we can not identify any specific mechanism (e.g.~leader-follower) of social information transfer.  But, in a generic way, our model assumes that migration is driven by a collective decision for the timing and locations of seasonal ranges, consistent with the known social and exogenous (e.g.~daylength related) triggers for migration. Further, the underlying assumption of the migration ``urge'' is consistent with the strong endogenous programs to migrate, e.g.~the seasonal restlessness known as \emph{Zugunruhe} exhibited by many birds \citep{Berthold1999, Helm2006}.  However, in its generic diffusion-based approach to randomness, our model indirectly captures individual-level variation in migration parameters, an inevitable property of any population-level process \citep{Gurarie2019}.

@@

	Given the slow scale of fitness selection and the constant change in environmental conditions, it is possible that certain inherent properties of populations, for example the ``conservatism'' captured by the $\kappa$ parameter, are themselves selected for to maximize resilience over a long time scale in stochastic environments.  The structure of the reference memory in our model was a rather simplistic approach to introduce conservatism or lag to the shifting migration parameters. In our model that reference memory is eventually entirely forgotten, whereas a more sophisticated approach would separate a slowly varying cultural memory, perhaps that is transmitted genetically or culturally, i.e.~on the scale of generations, against shorter-scaled responses.  In an evolutionary model, we might hypothesize that the overall rates of long- and short-term memory shifts would be related both to the scales of short and long-term fluctuation of the resource, i.e.~the auto-correlation scale, strength of trends, and stochasticity of the resource dynamics.  \\

@@

	\subsection{Drifting resource}
	The drifting resource function has the following properties:
